---
title: "Smoking and Asthma Analysis"
author: "Group R: Hanif, Chuanbo, Meihe, Mingzhi, Yumeng"
date: "`r Sys.Date()`"
output: html_document
---

## Initialization & Setup

```{r initialization, include=FALSE}
# Set global chunk options: show code but hide messages and warnings
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# List of required packages
required_packages <- c("dplyr", "ggplot2", "broom", "tidyr", "srvyr", 
                       "survey", "ipumsr", "here", "car", "kableExtra", 
                       "modelsummary", "margins")

# Function to load or install packages if they're missing
load_or_install <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}

# Load all packages silently
invisible(lapply(required_packages, load_or_install))
```

## Load & Clean Data

```{r load_clean_data}
# Load dataset
ddi <- read_ipums_ddi("nhis_00012.xml")
data <- read_ipums_micro(ddi)

data_cleaned <- data %>%
  # Select only the relevant variables
  select(SMOKESTATUS2, ASTHMAEV, CIGSDAY2, AGE, SEX, EDUC, GOTWAGE, REGION) %>%
  # Remove rows missing key smoking or asthma information
  filter(!is.na(SMOKESTATUS2), !is.na(ASTHMAEV)) %>%
  mutate(
    # Clean cigarettes per day: set NIU codes to NA, keep 95 as is, convert others to numeric
    CIGSDAY2_cleaned = case_when(
      CIGSDAY2 %in% c(97, 98, 99) ~ NA_real_,
      CIGSDAY2 == 95 ~ 95,
      TRUE ~ as.numeric(CIGSDAY2)
    ),
    # Define smoking status based on SMOKESTATUS2 codes
    Smoking_Status = case_when(
      SMOKESTATUS2 %in% c(10, 11, 12, 13) ~ "Current Smoker",
      SMOKESTATUS2 == 20 ~ "Former Smoker",
      SMOKESTATUS2 == 30 ~ "Never Smoked",
      SMOKESTATUS2 %in% c(40, 90) ~ "Unknown",
      TRUE ~ "Unknown"
    ),
    # For current smokers, categorize intensity based on cigarettes per day
    Smoking_Intensity = case_when(
      Smoking_Status == "Current Smoker" & CIGSDAY2 <= 5 ~ "Current Smoker (Light)",
      Smoking_Status == "Current Smoker" & CIGSDAY2 <= 10 ~ "Current Smoker (Moderate)",
      Smoking_Status == "Current Smoker" & CIGSDAY2 > 10 ~ "Current Smoker (Heavy)",
      TRUE ~ Smoking_Status
    ),
    # Recode asthma variable: 1 indicates an asthma attack
    Asthma_Attack = if_else(ASTHMAEV == 2, 1, 0),
    # Recode age: set certain codes to NA or top-code to 85
    age = case_when(
      AGE %in% c(997, 998, 999) ~ NA_real_,
      AGE %in% c(085, 090, 099) ~ 85,
      TRUE ~ as.numeric(AGE)
    ),
    # Convert SEX to a factor with appropriate labels
    gender = factor(case_when(
      SEX == 1 ~ "Male",
      SEX == 2 ~ "Female",
      TRUE ~ NA_character_
    ), levels = c("Male", "Female")),
    # Recode education: group into "Bachelor+" or "No Bachelor"
    education = factor(case_when(
      EDUC %in% c(400, 501, 503, 505) ~ "Bachelor+",
      EDUC %in% c(100:116, 200:303) ~ "No Bachelor",
      TRUE ~ NA_character_
    ), levels = c("No Bachelor", "Bachelor+")),
    # Recode income based on wage status
    income = factor(case_when(
      GOTWAGE == 2 ~ "Yes",
      GOTWAGE == 1 ~ "No",
      TRUE ~ NA_character_
    ), levels = c("No", "Yes")),
    # Map region codes to names
    region = factor(case_when(
      REGION == 1 ~ "Northeast",
      REGION == 2 ~ "Midwest",
      REGION == 3 ~ "South",
      REGION == 4 ~ "West",
      TRUE ~ NA_character_
    ), levels = c("Northeast", "Midwest", "South", "West"))
  ) %>%
  # Remove observations missing any control variable
  filter(!is.na(age), !is.na(gender), !is.na(education), !is.na(income), !is.na(region)) %>%
  # Set reference categories for regression analyses
  mutate(
    Smoking_Intensity = relevel(factor(Smoking_Intensity), ref = "Never Smoked"),
    education = relevel(education, ref = "No Bachelor"),
    income = relevel(income, ref = "No"),
    region = relevel(region, ref = "Northeast")
  )
```



##Table experiment
```{r summary_table_with_dummies, message=FALSE, warning=FALSE}
library(dplyr)
library(kableExtra)

# 1. Add dummy columns for each smoking category
data_cleaned <- data_cleaned %>%
  mutate(
    # Make ASTHMAEV a 0/1 variable if needed:
   ASTHMAEV = if_else(ASTHMAEV == 2, 1, 0),
    
    # Create separate 0/1 columns for each smoking category
    `Never Smoked`    = if_else(Smoking_Intensity == "Never Smoked", 1, 0),
    `Former Smoker`   = if_else(Smoking_Intensity == "Former Smoker", 1, 0),
    `Light Smoker`    = if_else(Smoking_Intensity == "Current Smoker (Light)", 1, 0),
    `Moderate Smoker` = if_else(Smoking_Intensity == "Current Smoker (Moderate)", 1, 0),
    `Heavy Smoker`    = if_else(Smoking_Intensity == "Current Smoker (Heavy)", 1, 0),
    `Unknown`         = if_else(Smoking_Intensity == "Unknown", 1, 0)
  )

# 2. Define all variables + optional definitions
variable_info <- tibble::tribble(
  ~Variable,             ~Definition,
  "ASTHMAEV",            "1 if ever diagnosed w/ asthma, else 0",
  "Never Smoked",        "1 if never smoked, else 0",
  "Former Smoker",       "1 if formerly smoked, else 0",
  "Light Smoker",        "1 if currently a light smoker, else 0",
  "Moderate Smoker",     "1 if currently a moderate smoker, else 0",
  "Heavy Smoker",        "1 if currently a heavy smoker, else 0",
  "Unknown",             "1 if smoking status is unknown",
  "CIGSDAY2_cleaned",    "Number of cigarettes smoked per day",
  "AGE",                 "Age of individual",
  "SEX",                 "1 is male, 2 is female",
  "EDUC",                "1 is Bachelor+, 0 is no Bachelor, Education status)",
  "GOTWAGE",             "1 is employed, 0 is not employed, Employment status",
  "REGION",              "Region: 1.NE, 2.MW, 3.South, 4.West"
)

# 3. Define the function to get summary stats for one variable
get_summary_stats <- function(data, varname) {
  x <- data[[varname]]  
  obs <- sum(!is.na(x))
  
  if (is.numeric(x)) {
    the_mean <- round(mean(x, na.rm = TRUE), 2)
    the_sd   <- round(sd(x,   na.rm = TRUE), 2)
    the_min  <- round(min(x,  na.rm = TRUE), 2)
    the_max  <- round(max(x,  na.rm = TRUE), 2)
  } else {
    the_mean <- NA
    the_sd   <- NA
    the_min  <- NA
    the_max  <- NA
  }
  
  tibble(
    Obs  = obs,
    Mean = the_mean,
    SD   = the_sd,
    Min  = the_min,
    Max  = the_max
  )
}

# 4. Build the summary table for each variable in variable_info
final_summary <- variable_info %>%
  rowwise() %>%
  mutate(stats = list(get_summary_stats(data_cleaned, Variable))) %>%
  tidyr::unnest(cols = c(stats)) %>%
  ungroup()

# 5. Print with kable
kable_table <- final_summary %>%
  kable(
    format    = "html",
    caption   = "Basic Summary Statistics (With Dummy Columns)",
    col.names = c("Variable", "Definition", "Obs", "Mean", "SD", "Min", "Max")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE,
    position          = "center"
  )

print(kable_table)
```


## Descriptive Table: Demographics

```{r descriptive_table_demographics, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(kableExtra)

# Summarize key demographic variables
descriptive_demographics <- data_cleaned %>%
  group_by(Smoking_Intensity) %>%
  summarise(
    Observations      = n(),
    Mean_Age          = mean(age, na.rm = TRUE),
    SD_Age            = sd(age, na.rm = TRUE),
    Prop_Male         = mean(gender == "Male", na.rm = TRUE),
    Prop_Female       = mean(gender == "Female", na.rm = TRUE),
    Prop_BachelorPlus = mean(education == "Bachelor+", na.rm = TRUE),
    Prop_IncomeYes    = mean(income == "Yes", na.rm = TRUE),
    Prop_Asthma       = mean(Asthma_Attack == 1, na.rm = TRUE)
  ) %>%
  ungroup()

# Pivot so that Smoking_Intensity becomes rows, metrics become columns
descriptive_demographics_flipped <- descriptive_demographics %>%
  pivot_longer(
    cols = -Smoking_Intensity, 
    names_to = "Metric", 
    values_to = "Value"
  ) %>%
  mutate(
    Value = round(Value, 3),  # Round numeric values
    Metric = dplyr::case_when(
      Metric == "Observations"      ~ "Number of Observations",
      Metric == "Mean_Age"          ~ "Mean Age",
      Metric == "SD_Age"            ~ "SD of Age",
      Metric == "Prop_Male"         ~ "% Male",
      Metric == "Prop_Female"       ~ "% Female",
      Metric == "Prop_BachelorPlus" ~ "% with Bachelor+",
      Metric == "Prop_IncomeYes"    ~ "% with Income = Yes",
      Metric == "Prop_Asthma"       ~ "% with Asthma Attack",
      TRUE                          ~ Metric
    )
  ) %>%
  pivot_wider(
    names_from = Metric, 
    values_from = Value
  )

# Render the flipped table
descriptive_demographics_flipped %>%
  kable(
    format  = "html",
    caption = "Descriptive Characteristics by Smoking Intensity (Demographics)"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE,
    position          = "center"
  )
```

## Summary Table: Cigarette Metrics

```{r pivoted_metrics_table, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(kableExtra)

# (Optional) Re-factor Smoking_Intensity to ensure correct column order
data_cleaned <- data_cleaned %>%
  mutate(Smoking_Intensity = factor(
    Smoking_Intensity,
    levels = c("Never Smoked", "Former Smoker", "Unknown",
               "Current Smoker (Light)", "Current Smoker (Moderate)", "Current Smoker (Heavy)")
  ))

# Summarize cigarette-related metrics
combined_table <- data_cleaned %>%
  group_by(Smoking_Intensity) %>%
  summarise(
    N                 = n(),
    Asthma_No         = sum(Asthma_Attack == 0, na.rm = TRUE),
    Asthma_Yes        = sum(Asthma_Attack == 1, na.rm = TRUE),
    Mean_CIGSDAY      = mean(CIGSDAY2_cleaned, na.rm = TRUE),
    SD_CIGSDAY        = sd(CIGSDAY2_cleaned, na.rm = TRUE),
    Median_CIGSDAY    = median(CIGSDAY2_cleaned, na.rm = TRUE),
    IQR_CIGSDAY       = IQR(CIGSDAY2_cleaned, na.rm = TRUE),
    Min_CIGSDAY       = min(CIGSDAY2_cleaned, na.rm = TRUE),
    Max_CIGSDAY       = max(CIGSDAY2_cleaned, na.rm = TRUE),
    Percent_with_Asthma = 100 * mean(Asthma_Attack == 1, na.rm = TRUE),
    .groups = "drop"
  )

# Pivot the table
flipped_table <- combined_table %>%
  pivot_longer(
    cols = -Smoking_Intensity,
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(Metric = dplyr::recode(
    Metric,
    "N"                   = "Sample Size",
    "Asthma_No"           = "Count (No Asthma Attack)",
    "Asthma_Yes"          = "Count (Asthma Attack)",
    "Mean_CIGSDAY"        = "Mean Cigarettes/Day",
    "SD_CIGSDAY"          = "SD Cigarettes/Day",
    "Median_CIGSDAY"      = "Median Cigarettes/Day",
    "IQR_CIGSDAY"         = "Interquartile Range Cigarettes/Day",
    "Min_CIGSDAY"         = "Minimum Cigarettes/Day",
    "Max_CIGSDAY"         = "Maximum Cigarettes/Day",
    "Percent_with_Asthma" = "Percentage with Asthma Attack"
  )) %>%
  pivot_wider(names_from = Smoking_Intensity, values_from = Value) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# Same styling as the first table
final_table <- flipped_table %>%
  kable(
    format   = "html", 
    caption  = "Cigarette Metrics by Smoking Intensity",
    col.names = colnames(.)
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position   = "center"
  )

print(final_table)
```

## Visualization

```{r visualization, message=FALSE, warning=FALSE}
library(ggplot2)

# Bar Chart: Count of Asthma Attacks by Smoking Status
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "dodge") +
  labs(
    title = "Asthma Attack Count by Smoking Status",
    x = "Smoking Status", 
    y = "Count", 
    fill = "Asthma Attack"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("#1f78b4", "#33a02c"))

# Stacked Bar Chart: Proportion of Asthma Attacks by Smoking Status
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "fill") +
  labs(
    title = "Proportion of Asthma Attacks by Smoking Status",
    x = "Smoking Status", 
    y = "Proportion", 
    fill = "Asthma Attack"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("#1f78b4", "#33a02c"))
```

## Logistic Regression Model

```{r}
# Run logistic regression
logit_model <- glm(
  Asthma_Attack ~ Smoking_Intensity + gender + age + income + education + region, 
  data = data_cleaned, 
  family = binomial()
)

# Format with robust standard errors and p-values
modelsummary(
  list("Logistic Regression" = logit_model),
  exponentiate = TRUE, 
  stars = c('*' = .1, '**' = .05, '***' = .01),
  output = "html"
)
# To compare with the robust standard error
modelsummary(
  list("Logistic Regression" = logit_model),
  exponentiate = TRUE, 
  vcov = "robust",
  stars = c('*' = .1, '**' = .05, '***' = .01),
  output = "html"
)

## Bivariate Test: Chi-Square
chi_test <- chisq.test(table(data_cleaned$Smoking_Status, data_cleaned$Asthma_Attack))
chi_test
```

```{r}
# Load required packages
library(dplyr)
library(ggplot2)
library(modelsummary)
library(margins)
library(knitr)
library(kableExtra)

# --- Estimate Models ---
# LPM
m_lpm <- lm(Asthma_Attack ~ Smoking_Intensity + gender + age + income + education + region,
            data = data_cleaned)

# Compute predicted probabilities for later use in the histogram
data_cleaned$pred_lpm <- predict(m_lpm, data_cleaned)

# Logit model
m_logit <- glm(Asthma_Attack ~ Smoking_Intensity + gender + age + income + education + region,
               data = data_cleaned, family = binomial(link = "logit"))

# Probit model
m_probit <- glm(Asthma_Attack ~ Smoking_Intensity + gender + age + income + education + region,
                data = data_cleaned, family = binomial(link = "probit"))

# --- Regression Results: Estimates Only ---
# Convert regression output to a clean table
reg_table <- msummary(
  list("LPM" = m_lpm, "Logit" = m_logit, "Probit" = m_probit),
  vcov = rep("robust", 3),
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  output = "data.frame"
)

# Clean variable names to natural language
reg_table$term <- gsub("Smoking_Intensity", "", reg_table$term)
reg_table$term <- gsub("genderMale", "Male", reg_table$term)
reg_table$term <- gsub("genderFemale", "Female", reg_table$term)
reg_table$term <- gsub("educationBachelor\\+", "Bachelor's Degree or Higher", reg_table$term)
reg_table$term <- gsub("incomeYes", "Has Income", reg_table$term)
reg_table$term <- gsub("region", "Region: ", reg_table$term)

# Extract only coefficient estimates
estimates_table <- reg_table %>%
  filter(statistic == "estimate") %>%
  select(term, LPM, Logit, Probit)

# Display regression estimates table
cat("### Regression Coefficients\n")
kable(estimates_table, format = "html", caption = "Regression Coefficients (Robust SE)") %>%
  kable_styling(full_width = FALSE)

# --- Regression Results: Standard Errors ---
std_error_table <- reg_table %>%
  filter(statistic == "std.error") %>%
  select(term, LPM, Logit, Probit)

cat("\n### Standard Errors\n")
kable(std_error_table, format = "html", caption = "Standard Errors for Regression Coefficients") %>%
  kable_styling(full_width = FALSE)

# --- Predicted Probabilities Plot ---
cat("\n### Histogram of Predicted Probabilities (LPM)\n")
p <- ggplot(data_cleaned, aes(x = pred_lpm)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Predicted Probabilities (LPM)",
       x = "Predicted Probability", y = "Count")
print(p)
ggsave("predicted_probabilities.png", p, width = 8, height = 6, dpi = 300)

# --- Outside Bounds ---
outside_bounds <- mean(data_cleaned$pred_lpm < 0 | data_cleaned$pred_lpm > 1)
cat("\n**Fraction of predictions outside [0,1]:** ", round(outside_bounds, 3), "\n")

# --- Marginal Effects Table ---
# Calculate marginal effects for the logit model and display as a table.
marginal_effects <- margins(m_logit)
marginal_table <- as.data.frame(summary(marginal_effects))

cat("\n### Marginal Effects (Logit Model)\n")
kable(marginal_table, format = "html", caption = "Marginal Effects for Logit Model") %>%
  kable_styling(full_width = FALSE)
```

## Multicollinearity Test

```{r}
# Logistic 
model_interaction <- glm(Asthma_Attack ~ Smoking_Intensity * region + 
                         age + gender + education + income,
                         data = data_cleaned, family = binomial)

summary(model_interaction)
library(car)
vif_model <- glm(Asthma_Attack ~ Smoking_Intensity * region + age + gender + education + income,
                 data = data_cleaned, family = binomial)

vif(vif_model)  

model_no_interaction <- glm(Asthma_Attack ~ Smoking_Intensity + region + 
                            age + gender + education + income,
                            data = data_cleaned, family = binomial)

summary(model_no_interaction)
vif(model_no_interaction)
# We find that multicollinearity has been significantly reduced after change to the noninteraction model, which means the interaction term introduced multicollinearity, inflating standard errors 
```

## Robustness Checks

```{r}
# Probit regress
model_probit <- glm(Asthma_Attack ~ Smoking_Intensity + 
                    age + gender + education + income,
                    data = data_cleaned, family = binomial(link = "probit"))

summary(model_probit)

# LPM regress
model_lpm <- lm(Asthma_Attack ~ Smoking_Intensity + 
                age + gender + education + income,
                data = data_cleaned)

summary(model_lpm)

library(sandwich)
library(lmtest)

# Cluster SE
model_clustered <- glm(Asthma_Attack ~ Smoking_Intensity * region + 
                       age + gender + education + income,
                       data = data_cleaned, family = binomial)

coeftest(model_clustered, vcov = vcovCL, cluster = ~region)
```

## Heterogeneity Test

```{r}
# Define a function to run logistic regression and return ORs
run_logit_by_region <- function(region_name) {
  model <- glm(Asthma_Attack ~ Smoking_Intensity + age + gender + education + income, 
               data = data_cleaned %>% filter(region == region_name), 
               family = binomial)
  
  # Extract coefficients, standard errors, and p-values
  tidy_model <- tidy(model) %>%
    mutate(region = region_name,
           OR = exp(estimate),  # Convert to Odds Ratio
           OR_low = exp(estimate - 1.96 * std.error),  # 95% CI Lower Bound
           OR_high = exp(estimate + 1.96 * std.error), # 95% CI Upper Bound
           p_value = case_when(
             p.value < 0.001 ~ "***",
             p.value < 0.01 ~ "**",
             p.value < 0.05 ~ "*",
             p.value < 0.1 ~ ".",
             TRUE ~ ""
           ),
           OR_formatted = paste0(round(OR, 2), " [", 
                                 round(OR_low, 2), ", ", 
                                 round(OR_high, 2), "] ", p_value))
  
  return(tidy_model)
}

# Run logistic regressions for each region
regions <- c("Northeast", "Midwest", "South", "West")
results_list <- lapply(regions, run_logit_by_region)

# Combine results into a single dataframe
results_df <- bind_rows(results_list)

# Pivot the results to make a side-by-side comparison table
results_table <- results_df %>%
  select(term, region, OR_formatted) %>%
  pivot_wider(names_from = region, values_from = OR_formatted)

# Create a nicely formatted table using kableExtra
results_table %>%
  kable("html", caption = "Odds Ratios (OR) for Asthma Across Regions") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

## Heterogeneity Test for the sex

```{r}
# Define a function to run logistic regression and return ORs
run_logit_by_gender <- function(gender_name) {
  model <- glm(Asthma_Attack ~ Smoking_Intensity + age + education + income + region, 
               data = data_cleaned %>% filter(gender == gender_name), 
               family = binomial)
  
  # Extract coefficients, standard errors, and p-values
  tidy_model <- tidy(model) %>%
    mutate(gender = gender_name,
           OR = exp(estimate),  # Convert to Odds Ratio
           OR_low = exp(estimate - 1.96 * std.error),  # 95% CI Lower Bound
           OR_high = exp(estimate + 1.96 * std.error), # 95% CI Upper Bound
           p_value = case_when(
             p.value < 0.001 ~ "***",
             p.value < 0.01 ~ "**",
             p.value < 0.05 ~ "*",
             p.value < 0.1 ~ ".",
             TRUE ~ ""
           ),
           OR_formatted = paste0(round(OR, 2), " [", 
                                 round(OR_low, 2), ", ", 
                                 round(OR_high, 2), "] ", p_value))
  
  return(tidy_model)
}

# Run logistic regressions for Male and Female separately
genders <- c("Male", "Female")
results_list <- lapply(genders, run_logit_by_gender)

# Combine results into a single dataframe
results_df <- bind_rows(results_list)

# Pivot the results to make a side-by-side comparison table
results_table <- results_df %>%
  select(term, gender, OR_formatted) %>%
  pivot_wider(names_from = gender, values_from = OR_formatted)

# Create a nicely formatted table using kableExtra
results_table %>%
  kable("html", caption = "Odds Ratios (OR) for Asthma by Gender") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

## Residual Plot for Logistic Regression

```{r}
# Extract residuals and fitted values
residuals_data <- data.frame(
  Fitted = fitted(logit_model),
  Residuals = residuals(logit_model)
)

# Plot residuals
ggplot(residuals_data, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residual Plot: Logistic Regression", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
```

## Predicted Probabilities Distribution

```{r}
# Get predicted probabilities
data_cleaned$pred_prob <- predict(logit_model, type = "response")

# Plot histogram
ggplot(data_cleaned, aes(x = pred_prob)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Predicted Probabilities", x = "Predicted Probability", y = "Count")
```

## Marginal Effects for Logistic Regression

```{r}
# Compute marginal effects for logistic regression
marginal_effects <- margins(logit_model)

# Summarize Average Marginal Effects
summary(marginal_effects)
```
