---
title: "Smoking and Asthma Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r}
# Load necessary packages
target_packages <- c("dplyr", "ggplot2", "broom", "tidyr", "srvyr", "survey", "ipumsr", "here", "car", "kableExtra", "modelsummary", "margins")

package.check <- lapply(
  target_packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

## Load & Clean Data
```{r}
# Load dataset
ddi <- read_ipums_ddi("nhis_00010.xml")
data <- read_ipums_micro(ddi)

# Select relevant variables
data_cleaned <- data %>%
  select(SMOKESTATUS2, ASTHMAEV, CIGSDAY2) %>%
  filter(!is.na(SMOKESTATUS2) & !is.na(ASTHMAEV)) %>%
  mutate(
    # 定义吸烟状态
    Smoking_Status = case_when(
      SMOKESTATUS2 %in% c(10, 11, 12, 13) ~ "Current Smoker",
      SMOKESTATUS2 == 20 ~ "Former Smoker",
      SMOKESTATUS2 == 30 ~ "Never Smoked",
      SMOKESTATUS2 %in% c(40, 90) ~ "Unknown",
      TRUE ~ "Unknown"
    ),
    # 定义每日吸烟量分类（仅针对当前吸烟者）
    Smoking_Intensity = case_when(
      Smoking_Status == "Current Smoker" & CIGSDAY2 <= 5 ~ "Current Smoker (Light)",
      Smoking_Status == "Current Smoker" & CIGSDAY2 <= 10 ~ "Current Smoker (Moderate)",
      Smoking_Status == "Current Smoker" & CIGSDAY2 > 10 ~ "Current Smoker (Heavy)",
      TRUE ~ Smoking_Status  # 非当前吸烟者保持原状态
    ),
    Asthma_Attack = if_else(ASTHMAEV == 2, 1, 0)
  ) %>%
  # 将 Never Smoked 设为参考组
  mutate(
    Smoking_Intensity = factor(Smoking_Intensity),
    Smoking_Intensity = relevel(Smoking_Intensity, ref = "Never Smoked")
  )
```

## Summary Tables
```{r}
# Frequency Table
summary_table <- data_cleaned %>%
  group_by(Smoking_Status, Asthma_Attack) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Asthma_Attack, values_from = Count, names_prefix = "Asthma_") %>%
  kable(format = "html", caption = "Asthma Attack Cases by Smoking Status") %>%
  kable_styling()

print(summary_table)
```

## Visualization
```{r}
# Bar Chart: Asthma Attack Count by Smoking Status
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "dodge") +
  labs(title = "Asthma Attack Frequency by Smoking Status", x = "Smoking Status", y = "Count", fill = "Asthma Attack") +
  theme_minimal() +
  scale_fill_manual(values = c("#1f78b4", "#33a02c"))

# Stacked Bar Chart: Proportion of Asthma Attacks per Smoking Category
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Asthma Attacks by Smoking Status", x = "Smoking Status", y = "Proportion", fill = "Asthma Attack") +
  theme_minimal() +
  scale_fill_manual(values = c("#1f78b4", "#33a02c"))

# Facet Plot for Smoking Groups
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "fill") +
  facet_wrap(~ Smoking_Status) +
  labs(title = "Asthma Attack Proportions by Smoking Status", x = "Smoking Status", y = "Proportion") +
  theme_minimal()
```

## Statistical Analysis
```{r}
# Chi-Square Test
chi_test <- chisq.test(table(data_cleaned$Smoking_Status, data_cleaned$Asthma_Attack))
chi_test
```

## Logistic Regression Model
```{r}
# Run logistic regression
logit_model <- glm(
  Asthma_Attack ~ Smoking_Intensity, 
  data = data_cleaned, 
  family = binomial()
)

# Format with robust standard errors and p-values
modelsummary(
  list("Logistic Regression" = logit_model),
  exponentiate = TRUE, 
  stars = c('*' = .1, '**' = .05, '***' = .01),
  output = "html"
)
# To compare with the robust standard errorzlibrary(plm)
modelsummary(
  list("Logistic Regression" = logit_model),
  vcov = "robust",
  stars = c('*' = .1, '**' = .05, '***' = .01),
  output = "html"
)
```

## Residual Plot for Logistic Regression
```{r}
# Extract residuals and fitted values
residuals_data <- data.frame(
  Fitted = fitted(logit_model),
  Residuals = residuals(logit_model)
)

# Plot residuals
ggplot(residuals_data, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residual Plot: Logistic Regression", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
```

## Predicted Probabilities Distribution
```{r}
# Get predicted probabilities
data_cleaned$pred_prob <- predict(logit_model, type = "response")

# Plot histogram
ggplot(data_cleaned, aes(x = pred_prob)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Predicted Probabilities", x = "Predicted Probability", y = "Count")
```

## Marginal Effects for Logistic Regression
```{r}
# Compute marginal effects for logistic regression
marginal_effects <- margins(logit_model)

# Summarize Average Marginal Effects
summary(marginal_effects)
```

## Interpretation
The logistic regression results above provide insights into the relationship between smoking status and the likelihood of having an asthma attack in the past 12 months. The confidence intervals and significance levels indicate the strength of this relationship. Further investigation can include additional covariates to control for confounding factors.