---
title: "Smoking and Asthma Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r}
# Load necessary packages
target_packages <- c("dplyr", "ggplot2", "broom", "tidyr", "srvyr", 
"survey", "ipumsr", "here", "car", "kableExtra", "modelsummary", "margins")

# Check if packages are installed; if not, install them
package.check <- lapply(
  target_packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

## Load & Clean Data
```{r}
# Load dataset
ddi <- read_ipums_ddi("nhis_00012.xml")
data <- read_ipums_micro(ddi)

# Select relevant variables
data_cleaned <- data %>%
   select(SMOKESTATUS2, ASTHMAEV, CIGSDAY2, 
         AGE, SEX, EDUC, GOTWAGE, REGION) %>%
  filter(!is.na(SMOKESTATUS2) & !is.na(ASTHMAEV)) %>%
  mutate(
    CIGSDAY2_cleaned = case_when(
      CIGSDAY2 %in% c(97, 98, 99) ~ NA_real_,   # Keep these as NIU
      CIGSDAY2 == 95 ~ 95,                      # Keep 95
      TRUE ~ as.numeric(CIGSDAY2)               # Keep other number
    ),
    # Define smoking status
    Smoking_Status = case_when(
      SMOKESTATUS2 %in% c(10, 11, 12, 13) ~ "Current Smoker",
      SMOKESTATUS2 == 20 ~ "Former Smoker",
      SMOKESTATUS2 == 30 ~ "Never Smoked",
      SMOKESTATUS2 %in% c(40, 90) ~ "Unknown",
      TRUE ~ "Unknown"
    ),
    # Define the number of cigarette for the current smoker
    Smoking_Intensity = case_when(
      Smoking_Status == "Current Smoker" & CIGSDAY2 <= 5 ~ "Current Smoker (Light)",
      Smoking_Status == "Current Smoker" & CIGSDAY2 <= 10 ~ "Current Smoker (Moderate)",
      Smoking_Status == "Current Smoker" & CIGSDAY2 > 10 ~ "Current Smoker (Heavy)",
      TRUE ~ Smoking_Status  
    ),
    Asthma_Attack = if_else(ASTHMAEV == 2, 1, 0),
    # Age (AGE)
    age = case_when(
      AGE %in% c(997, 998, 999) ~ NA_real_,  # Set unknown ages to missing
      AGE %in% c(085, 090, 099) ~ 85,        # Recode top-coded ages (85+)
      TRUE ~ as.numeric(AGE)                 # Keep other ages as numeric
    ),
    
    # Gender (SEX)
    gender = factor(case_when(
      SEX == 1 ~ "Male",
      SEX == 2 ~ "Female",
      TRUE ~ NA_character_  # Set unknown genders to missing
    ), levels = c("Male", "Female")),
    
    # Education (EDU)
    education = factor(case_when(
      EDUC %in% c(400, 501, 503, 505) ~ "Bachelor+",  # Bachelor's degree or higher
      EDUC %in% c(100:116, 200:303) ~ "No Bachelor",  # No bachelor's degree
      TRUE ~ NA_character_  # Set unknown education to missing
    ), levels = c("No Bachelor", "Bachelor+")),
    
    # Income (GOTWAGE)
    income = factor(case_when(
      GOTWAGE == 2 ~ "Yes",  # Has wage income
      GOTWAGE == 1 ~ "No",   # No wage income
      TRUE ~ NA_character_   # Set unknown income to missing
    ), levels = c("No", "Yes")),
    
    # Region (REGION)
    region = factor(case_when(
      REGION == 1 ~ "Northeast",
      REGION == 2 ~ "Midwest",
      REGION == 3 ~ "South",
      REGION == 4 ~ "West",
      TRUE ~ NA_character_  # Set unknown regions to missing
    ), levels = c("Northeast", "Midwest", "South", "West"))
  ) %>%
    # Remove rows with missing control variables
  filter(!is.na(age) & !is.na(gender) & !is.na(education) & 
         !is.na(income) & !is.na(region)) %>%
   # Set reference groups
  mutate(
    Smoking_Intensity = relevel(factor(Smoking_Intensity), ref = "Never Smoked"),
    education = relevel(education, ref = "No Bachelor"),  # Reference: No bachelor's degree
    income = relevel(income, ref = "No"),                # Reference: No wage income
    region = relevel(region, ref = "Northeast")          # Reference: Northeast region
  )
```

## Summary Tables
```{r}
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

# 1. Ensure Smoking_Intensity is a factor with the desired ordering
data_cleaned <- data_cleaned %>%
  mutate(Smoking_Intensity = factor(Smoking_Intensity, 
                                    levels = c("Never Smoked", "Former Smoker", "Unknown",
                                               "Current Smoker (Light)", "Current Smoker (Moderate)", "Current Smoker (Heavy)")))

# 2. Create a combined summary table grouped by Smoking_Intensity,
#    excluding metrics that are mostly zeros/NA's (e.g. we remove CV_CIGSDAY)
combined_table <- data_cleaned %>%
  group_by(Smoking_Intensity) %>%
  summarise(
    N = n(),
    Asthma_No = sum(Asthma_Attack == 0, na.rm = TRUE),
    Asthma_Yes = sum(Asthma_Attack == 1, na.rm = TRUE),
    Mean_CIGSDAY = mean(CIGSDAY2_cleaned, na.rm = TRUE),
    SD_CIGSDAY = sd(CIGSDAY2_cleaned, na.rm = TRUE),
    Median_CIGSDAY = median(CIGSDAY2_cleaned, na.rm = TRUE),
    IQR_CIGSDAY = IQR(CIGSDAY2_cleaned, na.rm = TRUE),
    Min_CIGSDAY = min(CIGSDAY2_cleaned, na.rm = TRUE),
    Max_CIGSDAY = max(CIGSDAY2_cleaned, na.rm = TRUE),
    Percent_with_Asthma = mean(Asthma_Attack, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# 3. Flip the table so that summary metrics become rows and Smoking_Intensity groups become columns.
#    Then rename the raw metric names to reader-friendly labels.
flipped_table <- combined_table %>%
  pivot_longer(-Smoking_Intensity, names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = dplyr::recode(Metric,
                                "N" = "Sample Size",
                                "Asthma_No" = "Count (No Asthma Attack)",
                                "Asthma_Yes" = "Count (Asthma Attack)",
                                "Mean_CIGSDAY" = "Mean Cigarettes/Day",
                                "SD_CIGSDAY" = "SD Cigarettes/Day",
                                "Median_CIGSDAY" = "Median Cigarettes/Day",
                                "IQR_CIGSDAY" = "Interquartile Range Cigarettes/Day",
                                "Min_CIGSDAY" = "Minimum Cigarettes/Day",
                                "Max_CIGSDAY" = "Maximum Cigarettes/Day",
                                "Percent_with_Asthma" = "Percentage with Asthma Attack"
  )) %>%
  pivot_wider(names_from = Smoking_Intensity, values_from = Value) %>%
  # Round numeric columns to 2 decimal places for a neat output
  mutate(across(where(is.numeric), ~ round(., 2)))

# 4. Add a design touch with kableExtra.
#    Here we also sort the columns based on the factor levels we set earlier.
final_table <- flipped_table %>%
  kable(format = "html", caption = "Combined Summary Table by Smoking Intensity", 
        col.names = colnames(.)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, position = "center")

# Print the final table
print(final_table)
   
```


## Visualization
```{r}
# Bar Chart: Asthma Attack Count by Smoking Status
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "dodge") +
  labs(title = "Asthma Attack Frequency by Smoking Status", x = "Smoking Status", y = "Count", fill = "Asthma Attack") +
  theme_minimal() +
  scale_fill_manual(values = c("#1f78b4", "#33a02c"))

# Stacked Bar Chart: Proportion of Asthma Attacks per Smoking Category
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Asthma Attacks by Smoking Status", x = "Smoking Status", y = "Proportion", fill = "Asthma Attack") +
  theme_minimal() +
  scale_fill_manual(values = c("#1f78b4", "#33a02c"))

# Facet Plot for Smoking Groups
ggplot(data_cleaned, aes(x = Smoking_Status, fill = as.factor(Asthma_Attack))) +
  geom_bar(position = "fill") +
  facet_wrap(~ Smoking_Status) +
  labs(title = "Asthma Attack Proportions by Smoking Status", x = "Smoking Status", y = "Proportion") +
  theme_minimal()
```

## Statistical Analysis
```{r}
# Chi-Square Test
chi_test <- chisq.test(table(data_cleaned$Smoking_Status, data_cleaned$Asthma_Attack))
chi_test
```

## Logistic Regression Model
```{r}
# Run logistic regression
logit_model <- glm(
  Asthma_Attack ~ Smoking_Intensity + gender + age + income + education + region, 
  data = data_cleaned, 
  family = binomial()
)

# Format with robust standard errors and p-values
modelsummary(
  list("Logistic Regression" = logit_model),
  exponentiate = TRUE, 
  stars = c('*' = .1, '**' = .05, '***' = .01),
  output = "html"
)
# To compare with the robust standard error
modelsummary(
  list("Logistic Regression" = logit_model),
  exponentiate = TRUE, 
  vcov = "robust",
  stars = c('*' = .1, '**' = .05, '***' = .01),
  output = "html"
)
```
## Multicollinearity Test
```{r}
# Logistic 
model_interaction <- glm(Asthma_Attack ~ Smoking_Intensity * region + 
                         age + gender + education + income,
                         data = data_cleaned, family = binomial)

summary(model_interaction)
library(car)
vif_model <- glm(Asthma_Attack ~ Smoking_Intensity * region + age + gender + education + income,
                 data = data_cleaned, family = binomial)

vif(vif_model)  

model_no_interaction <- glm(Asthma_Attack ~ Smoking_Intensity + region + 
                            age + gender + education + income,
                            data = data_cleaned, family = binomial)

summary(model_no_interaction)
vif(model_no_interaction)
# We find that multicollinearity has been significantly reduced after change to the noninteraction model, which means the interaction term introduced multicollinearity, inflating standard errors 
```
## Robustness Checks
```{r}
# Probit regress
model_probit <- glm(Asthma_Attack ~ Smoking_Intensity + 
                    age + gender + education + income,
                    data = data_cleaned, family = binomial(link = "probit"))

summary(model_probit)

# LPM regress
model_lpm <- lm(Asthma_Attack ~ Smoking_Intensity + 
                age + gender + education + income,
                data = data_cleaned)

summary(model_lpm)

library(sandwich)
library(lmtest)

# Cluster SE
model_clustered <- glm(Asthma_Attack ~ Smoking_Intensity * region + 
                       age + gender + education + income,
                       data = data_cleaned, family = binomial)

coeftest(model_clustered, vcov = vcovCL, cluster = ~region)
```
## Heterogeneity Test
```{r}
# Define a function to run logistic regression and return ORs
run_logit_by_region <- function(region_name) {
  model <- glm(Asthma_Attack ~ Smoking_Intensity + age + gender + education + income, 
               data = data_cleaned %>% filter(region == region_name), 
               family = binomial)
  
  # Extract coefficients, standard errors, and p-values
  tidy_model <- tidy(model) %>%
    mutate(region = region_name,
           OR = exp(estimate),  # Convert to Odds Ratio
           OR_low = exp(estimate - 1.96 * std.error),  # 95% CI Lower Bound
           OR_high = exp(estimate + 1.96 * std.error), # 95% CI Upper Bound
           p_value = case_when(
             p.value < 0.001 ~ "***",
             p.value < 0.01 ~ "**",
             p.value < 0.05 ~ "*",
             p.value < 0.1 ~ ".",
             TRUE ~ ""
           ),
           OR_formatted = paste0(round(OR, 2), " [", 
                                 round(OR_low, 2), ", ", 
                                 round(OR_high, 2), "] ", p_value))
  
  return(tidy_model)
}

# Run logistic regressions for each region
regions <- c("Northeast", "Midwest", "South", "West")
results_list <- lapply(regions, run_logit_by_region)

# Combine results into a single dataframe
results_df <- bind_rows(results_list)

# Pivot the results to make a side-by-side comparison table
results_table <- results_df %>%
  select(term, region, OR_formatted) %>%
  pivot_wider(names_from = region, values_from = OR_formatted)

# Create a nicely formatted table using kableExtra
results_table %>%
  kable("html", caption = "Odds Ratios (OR) for Asthma Across Regions") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
## Heterogeneity Test for the sex
```{r}
# Define a function to run logistic regression and return ORs
run_logit_by_gender <- function(gender_name) {
  model <- glm(Asthma_Attack ~ Smoking_Intensity + age + education + income + region, 
               data = data_cleaned %>% filter(gender == gender_name), 
               family = binomial)
  
  # Extract coefficients, standard errors, and p-values
  tidy_model <- tidy(model) %>%
    mutate(gender = gender_name,
           OR = exp(estimate),  # Convert to Odds Ratio
           OR_low = exp(estimate - 1.96 * std.error),  # 95% CI Lower Bound
           OR_high = exp(estimate + 1.96 * std.error), # 95% CI Upper Bound
           p_value = case_when(
             p.value < 0.001 ~ "***",
             p.value < 0.01 ~ "**",
             p.value < 0.05 ~ "*",
             p.value < 0.1 ~ ".",
             TRUE ~ ""
           ),
           OR_formatted = paste0(round(OR, 2), " [", 
                                 round(OR_low, 2), ", ", 
                                 round(OR_high, 2), "] ", p_value))
  
  return(tidy_model)
}

# Run logistic regressions for Male and Female separately
genders <- c("Male", "Female")
results_list <- lapply(genders, run_logit_by_gender)

# Combine results into a single dataframe
results_df <- bind_rows(results_list)

# Pivot the results to make a side-by-side comparison table
results_table <- results_df %>%
  select(term, gender, OR_formatted) %>%
  pivot_wider(names_from = gender, values_from = OR_formatted)

# Create a nicely formatted table using kableExtra
results_table %>%
  kable("html", caption = "Odds Ratios (OR) for Asthma by Gender") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```
## Residual Plot for Logistic Regression
```{r}
# Extract residuals and fitted values
residuals_data <- data.frame(
  Fitted = fitted(logit_model),
  Residuals = residuals(logit_model)
)

# Plot residuals
ggplot(residuals_data, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residual Plot: Logistic Regression", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
```

## Predicted Probabilities Distribution
```{r}
# Get predicted probabilities
data_cleaned$pred_prob <- predict(logit_model, type = "response")

# Plot histogram
ggplot(data_cleaned, aes(x = pred_prob)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Predicted Probabilities", x = "Predicted Probability", y = "Count")
```

## Marginal Effects for Logistic Regression
```{r}
# Compute marginal effects for logistic regression
marginal_effects <- margins(logit_model)

# Summarize Average Marginal Effects
summary(marginal_effects)
```

## Interpretation
The logistic regression results above provide insights into the relationship between smoking status and the likelihood of having an asthma attack in the past 12 months. The confidence intervals and significance levels indicate the strength of this relationship. Further investigation can include additional covariates to control for confounding factors.